Code for Breast Cancer Pain Study

### 2020-07-24 manuscript_code_20200722.Rmd
tables

```{r,echo=FALSE}
respvar = "Pain"
explvars = c("Pain","Poverty","Vacancy",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo","Surgery",
             "Race","Age",
             "Histology",
             "Primary.Site.1","Primary.Site.2",
             "HER2","ER","PR","Metastasis")

@@

@@

```{r,echo=FALSE}
p.glm = glm(Pain ~ Poverty, data = final.data, family = "poisson")
# summary(p.glm)
fhat = function(x) {
  exp(p.glm$coefficients[1]+p.glm$coefficients[2]*x)
}

@@

```{r,echo=FALSE}
p.glm = glm(Pain ~ Vacancy, data = final.data, family = "poisson")
# summary(p.glm)
fhat = function(x) {
  exp(p.glm$coefficients[1]+p.glm$coefficients[2]*x)
}

@@

```{r,echo=FALSE}
p.glm = glm(Pain ~ Vacancy, data = cdmd.pvt, family = "poisson")
# summary(p.glm)
# summary(zeroinfl(Pain ~ Poverty, data = cdmd.pvt, dist="poisson"))
fhat = function(x) {
  exp(p.glm$coefficients[1]+p.glm$coefficients[2]*x)
}

@@

```{r,echo=FALSE}
xlist = c("Poverty.f","marital.status.f","Primary.Payer.f2","Usual.Occupation.f2","race.f","stage.f","age.f","chemo.f","surgery.f")

@@

x = "Stage"
tmp <-
  cdmd.pvt %>%
  select("Pain","stage.f")
colnames(tmp) = c("Pain",x)
my_comparisons <- list( c("1", "2"), c("2", "3"), c("1", "3") )
base <- ggboxplot(tmp, x=x, y="Pain",
                  color = x, palette="jco",legend="none") +
  labs(x = NULL, y = "Pain", title = x) + 
  font("title",size=20,color="black") + 
  font("x.text", size = 20, color = "black") +
  font("ylab", size = 20, color = "black") + 
  font("y.text", size = 20, color = "black") +
  stat_compare_means(size=7, aes(label = "p.format"),show.legend=F,hide.ns=T,
                     label.y = 15) +
  stat_compare_means(comparisons = my_comparisons, aes(label = ..p.signif..),hide.ns=T)
p <- base +
  theme(plot.title=element_text(face="bold",size=20,margin=margin(t=10,b=20)))
ggsave(p, file=paste0(x,".pdf"),width=6, height=6)

@@

```{r,echo=FALSE,eval=FALSE}

>>>>

```{r,echo=FALSE,eval=FALSE}
respvar = "Pain"
explvars = c("Pain","Poverty","Vacancy",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo","Surgery",
             "Race","Age",
             "Histology",
             "Primary.Site.1","Primary.Site.2",
             "HER2","ER","PR","Metastasis")

@@

```{r,echo=FALSE}
explvars = c("Pain","Poverty","Vacancy",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo","Surgery",
             "Race","Age",
             "Histology",
             "Primary.Site.1","Primary.Site.2",
             "HER2","ER","PR","Metastasis")

get_summary = function(j) {
  x = final.data[,j]
  cname = colnames(final.data)[j]
  if (class(x)=="numeric") {
    tmp.sum = t(t(c(paste0(round(mean(x),digits=2)," (",round(sd(x),digits=2),")"),
                    paste0(round(median(x),digits=2)," [",round(min(x),digits=2),",",round(max(x),digits=2),"]"))))
    rownames(tmp.sum) = paste(cname,c("Mean (SD)","Median [Min,Max]"),sep="_")
  } else {
    tmp.sum = t(t(summary(x)))
    tmp.sum[,1] = paste(tmp.sum[,1],paste0("(",round(100*(tmp.sum[,1]/sum(tmp.sum[,1])),digits=1),"%)"))
    rownames(tmp.sum) = paste(cname,rownames(tmp.sum),sep="_")
  }
  return(tmp.sum)
}
table.sum1 = do.call("rbind",sapply(1:ncol(final.data),get_summary))

get_mean = function(j) {
  x = final.data[,j]
  cname = colnames(final.data)[j]
  if (class(x)=="numeric") {
    tmp.sum = t(t(c(NA,NA)))
  } else {
    tmp.sum = t(t(sapply(split(final.data[,"Pain"],final.data[,j]),function(x) 
  paste0(round(mean(x),digits=2)," [",paste0(round(quantile(x,c(0.25,0.75)),digits=2),collapse=","),"]"))))
  }
  return(tmp.sum)
}

get_pval = function(j) {
  x = final.data[,j]
  cname = colnames(final.data)[j]
  if (class(x)=="numeric") {
    tmp.sum = t(t(c(1,NA)))
  } else {
    tmp.sum = t(t(rep(NA,length(levels(x)))))
    tmp.sum[1,1] = 1
  }
  return(tmp.sum)
}

table.sum1 = cbind(do.call("rbind",sapply(1:ncol(final.data),get_summary)),
                   do.call("rbind",sapply(1:ncol(final.data),get_mean)),
                   do.call("rbind",sapply(1:ncol(final.data),get_pval)))
colnames(table.sum1) = c("No. (%)","Pain, mean [Q1,Q3]","P-value")

## htmlTable
library(htmlTable)

rowlabels = sapply(rownames(table.sum1),function(x) unlist(strsplit(x,split="_"))[1])
htmlTable(table.sum1,
          rgroup = c(gsub(pattern="[.]",replacement=" ",x=explvars[1:length(explvars)])),
          rnames = sapply(rownames(table.sum1),function(x) unlist(strsplit(x,split="_"))[2]),
          n.rgroup = sapply(explvars, function(x) length(which(rowlabels==x))),
          caption = "Table 1. Baseline characteristics, mean and interquartile range of pain intensity, nonparametric rank test")

```{r,echo=FALSE,eval=FALSE}
library(stats)
j = 4
colnames(final.data)[j]
kruskal.test(final.data[,"Pain"] ~ final.data[,j])
kruskal.test(final.data[-which(final.data[,j]=="Unknown"),"Pain"] ~ final.data[-which(final.data[,j]=="Unknown"),j])
å
kruskal.test(final.data[which(final.data[,j] %in% c("Employed","Unemployed")),"Pain"] ~ final.data[which(final.data[,j] %in% c("Employed","Unemployed")),j])

wilcox.test(final.data[-which(final.data[,j]=="Unknown"),"Pain"] ~ final.data[-which(final.data[,j]=="Unknown"),j])

compare_means(Pain ~ Occupation,data=final.data)
```

@@

```{r,echo=FALSE,eval=FALSE}
p.glm = glm(Pain ~ Poverty, data = final.data, family = "poisson")
# summary(p.glm)
fhat = function(x) {
  exp(p.glm$coefficients[1]+p.glm$coefficients[2]*x)
}

@@

```{r,echo=FALSE,eval=FALSE}
p.glm = glm(Pain ~ Vacancy, data = final.data, family = "poisson")
# summary(p.glm)
fhat = function(x) {
  exp(p.glm$coefficients[1]+p.glm$coefficients[2]*x)
}

@@

```{r,echo=FALSE,eval=FALSE}
p.glm = glm(Pain ~ Vacancy, data = cdmd.pvt, family = "poisson")
# summary(p.glm)
# summary(zeroinfl(Pain ~ Poverty, data = cdmd.pvt, dist="poisson"))
fhat = function(x) {
  exp(p.glm$coefficients[1]+p.glm$coefficients[2]*x)
}

@@

```{r,echo=FALSE,eval=FALSE}
xlist = c("Poverty.f","marital.status.f","Primary.Payer.f2","Usual.Occupation.f2","race.f","stage.f","age.f","chemo.f","surgery.f")

@@

x = "Stage"
tmp <-
  cdmd.pvt %>%
  select("Pain","stage.f")
colnames(tmp) = c("Pain",x)
my_comparisons <- list( c("1", "2"), c("2", "3"), c("1", "3") )
base <- ggboxplot(tmp, x=x, y="Pain",
                  color = x, palette="jco",legend="none") +
  labs(x = NULL, y = "Pain", title = x) + 
  font("title",size=20,color="black") + 
  font("x.text", size = 20, color = "black") +
  font("ylab", size = 20, color = "black") + 
  font("y.text", size = 20, color = "black") +
  stat_compare_means(size=7, aes(label = "p.format"),show.legend=F,hide.ns=T,
                     label.y = 15) +
  stat_compare_means(comparisons = my_comparisons, aes(label = ..p.signif..),hide.ns=T)
p <- base +
  theme(plot.title=element_text(face="bold",size=20,margin=margin(t=10,b=20)))
ggsave(p, file=paste0(x,".pdf"),width=6, height=6)
```

@@

```{r,echo=FALSE}
